---
title: "Problem Set #4"
author: "Nicholas Dow"
date: "February 27, 2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(readr)
library(gt)
library(ggplot2)
```
##Question 1
```{r madlibs, echo=FALSE, message=FALSE}
elections <- read_csv(file = "ps_4_elections-poll-nc09-3.csv")

madlib1 <- elections %>% 
  filter(response == "Dem") %>% 
  count()

fav_rep <- elections %>% 
  filter(hrep_fav == "Favorable") %>% 
  count()
und <- elections %>% 
  filter(response == "Und") %>% 
  count()
madlib2 <- fav_rep$n - und$n

madlib3 <- elections %>% 
  filter(gender != gender_combined) %>% 
  count()
  
madlib4 <- elections %>% 
  filter(race_eth == "White") %>% 
  filter(file_race_black != "White") %>% 
  count()

dem_time <- elections %>% 
  filter(response == "Dem") %>% 
  pull(timestamp)

rep_t <- elections %>% 
  filter(response == "Rep") %>% 
  pull(timestamp)

madlib5 <-round(rep_t[1] - dem_time[1])
```
There were `r madlib1` respondents who supported the Democratic candidate.

There were `r madlib2` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r madlib3` individuals for whom these variables have different values.

There are `r madlib4` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Dem came `r madlib5` minutes (rounded to the nearest minute) before the first response of Rep.

##Question 2

```{r echo=FALSE, warning=FALSE}

elections %>% 
  select(response, race_eth, final_weight) %>% 
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>%
  mutate(race_eth = fct_relevel(race_eth, "White", "Black", "Hispanic", "Asian", "Other")) %>%
  group_by(response,race_eth) %>% 
  summarize(all = sum(final_weight)) %>% 
  spread(key =  response, value = all, fill = 0) %>% 
  mutate(all = Dem + Rep + Und + `3`) %>% 
  mutate(Dem = Dem / all) %>% 
  mutate(Rep = Rep / all) %>% 
  mutate(Und = Und / all) %>% 
  select(-all,-`3`) %>% 
  gt() %>% 
  tab_header(
    title = "Polling Results by Race in North Carolina's 9th Congressional District") %>% 
  cols_label(
    race_eth = "",
    Dem = "DEM.",
    Rep = "REP.",
    Und = "UND."
  ) %>%
  tab_source_note(source_note = "Source: New York Times Upshot/Siena College 2018 live polls") %>% 
  fmt_percent(columns = vars(Dem, Rep, Und),
              decimals = 0) %>% 
  
  na_if(0) %>%
  fmt_missing(columns = vars(Und), rows = 4)
```

##Question 3

```{r echo=FALSE, warning=FALSE}
graphic <- elections %>% 
  select(educ, final_weight) %>% 
  filter(educ != "[DO NOT READ] Refused") %>% 
  mutate(educ = fct_relevel(educ, "Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree")) %>% 
  ggplot(aes(x = educ, y = final_weight)) +
  geom_violin() + 
  geom_jitter(width = 0.2, size = 1.25, alpha = 0.4) + 
  coord_flip() +
  labs(title = "The More Educated Matter Less in North Carolina 9th ",
       subtitle = "Poll gives more weight to people who are less likely to participate in polls",
       caption = "New York Times Upshot/Siena College 2018 live polls") +
  ylab("Weight Given to Respondent in Calculating Poll Results") +
  xlab(NULL)
graphic
```

##Question 4

```{r echo=FALSE, warning=FALSE}
resp_age <- elections %>% filter(response != "3") %>%  count(response, file_race, gender)
p <- ggplot(resp_age, aes(x = response, y = n, fill=file_race))
p + geom_col() +
  labs(title = "Gender Race Breakdown by Party", x = NULL, y = "Number of responses", fill = "Race") +
  facet_grid(~ gender) +
  coord_flip()
```