---
title: "Problem Set #4"
author: "Nicholas Dow"
date: "February 27, 2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)

#Tidyverse for data management

library(readr)

#reader to read in the csv file

library(gt)

#gt for table display

library(ggplot2)

#ggplot for visualization

```
##Question 1
```{r madlibs, echo=FALSE, message=FALSE}
elections <- read_csv(file = "ps_4_elections-poll-nc09-3.csv")

#Reading in csv file

madlib1 <- elections %>% 
  filter(response == "Dem") %>% 
  count()

#Just counting up the # of Dem respondents

fav_rep <- elections %>% 
  filter(hrep_fav == "Favorable") %>% 
  count()

#Counting # of people who view republicans favorably, saving to number, maybe could have done a clever summarise fun

und <- elections %>% 
  filter(response == "Und") %>% 
  count()

#Counting # undecided like above

madlib2 <- fav_rep$n - und$n

#Taking Diff of two values

madlib3 <- elections %>% 
  filter(gender != gender_combined) %>% 
  count()

#Counting where gender is unequal in 2 specified cols
  
madlib4 <- elections %>% 
  filter(race_eth == "White") %>% 
  filter(file_race_black != "White") %>% 
  count()

#Filtering and counting for conditions of question

dem_time <- elections %>% 
  filter(response == "Dem") %>% 
  pull(timestamp)

#Pull sorts our data/timestamp types for us, so we can use that later

rep_t <- elections %>% 
  filter(response == "Rep") %>% 
  pull(timestamp)

#we do *above* for reps

madlib5 <-round(rep_t[1] - dem_time[1])

#We take the diff of the top of the sorted lists to get the difference in amount of time

#We will put the madlibs in to the madlib below
```
There were `r madlib1` respondents who supported the Democratic candidate.

There were `r madlib2` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r madlib3` individuals for whom these variables have different values.

There are `r madlib4` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Dem came `r madlib5` minutes (rounded to the nearest minute) before the first response of Rep.

##Question 2

```{r echo=FALSE, warning=FALSE}

elections %>% 
  select(response, race_eth, final_weight) %>% 
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>%
  
  #We sort out the data that we need for processing
  
  mutate(race_eth = fct_relevel(race_eth, "White", "Black", "Hispanic", "Asian", "Other")) %>%
  
  #Create factors for the ordering of the list
  
  group_by(response,race_eth) %>% 
  summarize(all = sum(final_weight)) %>% 
  
  #Two Commands above are a cleaver way to sum up the final weights associated with resp eth combo
  
  spread(key =  response, value = all, fill = 0) %>% 
  
  #Make left side of rows response instead of having as a col and switches the other values to stay lined up with their combo from cols
  mutate(all = Dem + Rep + Und + `3`) %>% 
  mutate(Dem = Dem / all) %>% 
  mutate(Rep = Rep / all) %>% 
  mutate(Und = Und / all) %>% 
  select(-all,-`3`) %>% 
  
  #Makes a couple new cols from the ones provided, then removes the two uneeded cols
  
  gt() %>% 
  tab_header(
    title = "Polling Results by Race in North Carolina's 9th Congressional District") %>% 
  cols_label(
    race_eth = "",
    Dem = "DEM.",
    Rep = "REP.",
    Und = "UND."
  ) %>%
  tab_source_note(source_note = "Source: New York Times Upshot/Siena College 2018 live polls") %>% 
  
  #Gt puts into formatted table which can be customized with some of the other commands above. We give the table everything it needs to be an informative table
  fmt_percent(columns = vars(Dem, Rep, Und),
              decimals = 0) %>% 
  na_if(0) %>%
  fmt_missing(columns = vars(Und), rows = 4)

  #We format the values to be percents, or dashed

```

##Question 3

```{r echo=FALSE, warning=FALSE}
graphic <- elections %>% 
  select(educ, final_weight) %>% 
  filter(educ != "[DO NOT READ] Refused") %>% 
  
  #We sort out the data that we will be working with
  
  mutate(educ = fct_relevel(educ, "Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree")) %>% 
  
  #We turn educ into a factor in a specific order, so it shows up right later
  
  ggplot(aes(x = educ, y = final_weight)) +
  geom_violin() + 
  geom_jitter(width = 0.2, size = 1.25, alpha = 0.4) + 
  
  #We put our data into its visualizaiton, and specify how parts of the data should show up.
  
  coord_flip() +
  labs(title = "The More Educated Matter Less in North Carolina 9th ",
       subtitle = "Poll gives more weight to people who are less likely to participate in polls",
       caption = "New York Times Upshot/Siena College 2018 live polls") +
  ylab("Weight Given to Respondent in Calculating Poll Results") +
  xlab(NULL)

  #We Flip the coords and then label every part of the graph to make it informative
graphic
```

##Question 4

```{r echo=FALSE, warning=FALSE}
resp_race_age <- elections %>% 
  filter(response != "3") %>%  
  count(response, file_race, gender)

#We sort out the information we need, I couldn't find anything particularly interesting, so I picked out data based on the only relations I could find
#I count out the number of people in each catagory, I know that we have used final weights up to this point, but I don't use anything that I don't fully understand on principle.

p <- ggplot(resp_race_age, aes(x = response, y = n, fill=file_race))

#Specify the information and how it is being put into the visualization

p + geom_col() +
  labs(title = "Gender Race Breakdown by Party", x = NULL, y = "Number of responses", fill = "Race") +
  
  #Put the data into the visualization and give it the proper labels to make it better at conveying information.
  #Some people may not think the colors look good, I avoided changing the colors because fill was race and I'm not touching that with a 20ft pool. Dock me points I don't care, I'm not going to mess with color of the bars here.
  #It went into fill because there are a greater # of races vs gender, and the information can be communicated more efficently in this format.
  #We didn't split up race into multiple bars because some values were so small that the bars weren't easy to see. this method gives a decent relative size comparison by moving the data closer together

  facet_grid(~ gender) +
  
  #We facet_grid by gender to get two graphs to group information on gender to communicate some trends in gender.
  coord_flip()
#flips coords for labels to show up more cleanly
```